---
title: "ajz_esm206_lab5"
author: "AJ Zekanoski"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives 

- Getting counts for different groups
- Use {lubridate} to parse dates
- Find confidence intervals & do t-tests with t.test()
- Heat map with geom_tile()

```{r, include = FALSE}
library(tidyverse)
library(here)
library(janitor)
```

```{r}
# Import Lobster data

lobster_abundance <- read_csv(here::here("data", "lobster_abundance.csv"),
                              na = "-99999") %>% 
  clean_names()
```

Use the tidyr uncount() to expand a frequency variable into a bunch of observations

```{r}
lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)
```

Yay now each lobster has its own row!

Lets look at this a bunch of ways:
```{r}
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site),
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

ggplot(data = lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) +
  facet_wrap(~site, scales = "free")


```

use 'lubridate' to parse dates and times:

I'm going to add a new column wuith mutate that conatins my date as an actual date


```{r}
lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
```

Parse 'Date' column to get different pieces like month and year seperated

```{r}
lobster_parse_date <- lobster_date %>% 
  mutate(obs_month = lubridate::month(date_new, label = TRUE),
         obs_year = lubridate::year(date_new))
```

Count lobsters by different groupings...

Let's say I want to count the number of lobsters by year and month. count() does 3 things, group by, summarise, and find the length. then it ungroups. Real awesome for cat data.

```{r}
lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

lobster_ym

lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)

lobster_ysite

lobster_site <- lobster_parse_date %>% 
  count(site)

lobster_site

# group_by + summarize +n ()

lobster_summary <- lobster_parse_date %>% 
  group_by(site) %>% 
  summarise(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n()
  )

```

Confidence intervals and t-tests

Use the t-test() to fund confidence intervals and perform t-tests

```{r}
ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters)

```

Two-sample t-tests:
Is there a significant difference in mean lobster lenghths between Naples Reef and Mohawk Reef?

```{r}
napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample, mohk_sample)

mohk_napl_ttest
```



Maybe an easier way...
```{r}
lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)

```

```{r}
ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))
```

